\documentclass{ximera}

\input{../preamble.tex}
\title[Prerequisite Videos: ]{Algebra: Inequalities}
\author{Ben Sencindiver}

\outcome{Understand how to find the solutions to inequalities.}

\begin{document}
\begin{abstract}
  In this series of videos, we will explore how to find the solutions
  to algebraic inequalities.
\end{abstract}
\maketitle

The following videos talk about how to deal with solving inequalities:


%% Introduction Video
\textbf{Introduction to Algebraic Inequalities: What manipulations
can
be made and how they affect the solution}
%% Labeling this expandable option
\begin{flushright}
{\color{blue}(\emph{Click the arrow to the right to see the Introduction video.})}
\end{flushright}
\begin{center}
\begin{expandable}
\youtube{DpH2zd2ukaw}
\end{expandable}
\end{center}





%% Question 1
\textbf{Question 1: Basic linear inequalities 1}
\begin{question}
%% Labeling this expandable option
\begin{flushright}
{\color{blue}(\emph{Click the arrow to the right to see the first question.})}
\end{flushright}
\begin{center}
\begin{expandable}
\youtube{FGHe6wk5hV4}
%% Multiple Choice Question 1
{\color{blue}(\emph{Click the arrow to the right to see the question
posed at the end of the video.})}
\begin{expandable}
For what values of $x$ is it true that $-4x-7 <13$?\\
\begin{prompt}
The inequality is satisfied when $x$ is \wordChoice{\choice[correct]{greater than}
\choice{greater than or equal to}\choice{less than}
\choice{less than or equal to}} $\answer[given]{-5}$
\end{prompt}
%% Example 1
\begin{flushright}
{\color{blue}(\emph{Click the arrow to the right to see an example.})}
\end{flushright}
\begin{expandable}
Example 1
\youtube{Cdt0zB0h5xM}
\end{expandable}
\end{expandable}
\end{expandable}
\end{center}
\end{question}



%% Question 2
\textbf{Question 2: Basic linear inequalities 2}
\begin{question}
%% Labeling this expandable option
\begin{flushright}
{\color{blue}(\emph{Click the arrow to the right to see the definition
of trigonometric functions and second question.})}
\end{flushright}
\begin{center}
\begin{expandable}
\youtube{i2Gbziuo3mY}
%% Multiple Choice Question 2
{\color{blue}(\emph{Click the arrow to the right to see the question
posed at the end of the video.})}
\begin{expandable}
When is $-x+20\leq 8x+7$?\\
\begin{prompt}
The inequality is satisfied when $x$ is \wordChoice{\choice{greater than}
\choice[correct]{greater than or equal to}\choice{less than}
\choice{less than or equal to}} $\answer[given]{\frac{13}{9}}$
\end{prompt}
%% Example 2
\begin{flushright}
{\color{blue}(\emph{Click the arrow to the right to see an example.})}
\end{flushright}
\begin{expandable}
Example 2
\youtube{kZpUg-QxS1Q}
\end{expandable}
\end{expandable}
\end{expandable}
\end{center}
\end{question}



\end{document}
