\documentclass{ximera}

\input{../preamble.tex}
\title[Prerequisite Videos: ]{Function Notation}
\author{Ben Sencindiver}

\outcome{Understand common topics about function
notation, such as evaluation, domain, and range.}

\begin{document}
\begin{abstract}
  We go through common prerequisite topics related
  function notation, such as evaluation, domain,
  and range of functions.
\end{abstract}
\maketitle

The following videos will cover topics about Function Notation:




%% Question 3
\textbf{Question 3: Domains of Functions Given by Graphs}
\begin{question}
%% Labeling this expandable option
\begin{flushright}
{\color{blue}(\emph{Click the arrow to the right to see the third question.})}
\end{flushright}
\begin{center}
\begin{expandable}
\youtube{42wkLf53444}

%% Multiple Choice Question 3
{\color{blue}(\emph{Click the arrow to the right to see the question
posed in the video.})}

\begin{expandable}

What is the domain of the function $h(x)= x^2+6x -7$?

\begin{multipleChoice}
\choice[correct]{All Real Numbers}
\choice{All Real Numbers except $x=1$}
\choice{$[1,7]$}
\choice{$(1,7)$}
\choice{$[0,\infty)$}
\choice{$(-\infty,0]$}
\end{multipleChoice}

%% Refresher 3
\begin{flushright}
{\color{blue}(\emph{Click the arrow to the right to see a short refresher.})}
\end{flushright}
\begin{expandable}
Refresher 3
\youtube{qTo5dJBwe98}
\end{expandable}
%% Example 3
\begin{flushright}
{\color{blue}(\emph{Click the arrow to the right to see an example.})}
\end{flushright}
\begin{expandable}
Example
\youtube{En0FJS9osqA}
\end{expandable}
\end{expandable}
\end{expandable}
\end{center}
\end{question}


%% Question 4
\textbf{Question 4: Ranges of Functions Given by Graphs}
\begin{question}
%% Labeling this expandable option
\begin{flushright}
{\color{blue}(\emph{Click the arrow to the right to see the fourth question.})}
\end{flushright}
\begin{center}
\begin{expandable}
\youtube{d_-WFfh4JUo}
%% Multiple Choice Question 4
{\color{blue}(\emph{Click the arrow to the right to see the  question
posed in the video.})}
\begin{expandable}

What is the range of the function $h(x)= x^2+6x -7$?

\begin{multipleChoice}
\choice{All Real Numbers}
\choice{All Real Numbers except $x=1$}
\choice{$[1,7]$}
\choice{$(1,7)$}
\choice{$[0,\infty)$}
\choice{$[-7,\infty)$}
\choice[correct]{$[-16,\infty)$}
\end{multipleChoice}
%% Refresher 4
\begin{flushright}
{\color{blue}(\emph{Click the arrow to the right to see a short refresher.})}
\end{flushright}
\begin{expandable}
Refresher 4
\youtube{WjAgjKlwzVs}
\end{expandable}
%% Example 4
\begin{flushright}
{\color{blue}(\emph{Click the arrow to the right to see an example.})}
\end{flushright}
\begin{expandable}
Example Part 1
\youtube{nUVz6fmt45c}\\

Example Part 2
\youtube{Khi_cgBqCw4}\\

Example Part 3
\youtube{wcM8nhtk__c}\\

Example Part 4
\youtube{ZJGUm-zhgGQ}
\end{expandable}
\end{expandable}
\end{expandable}
\end{center}
\end{question}


\end{document}

