\documentclass{ximera}

\input{../preamble.tex}

\title{Using limits to detect asymptotes}

\begin{document}

\begin{abstract}
%Stuff can go here later if we want!
\end{abstract}

\maketitle

\begin{sectionOutcomes}

After completing this section, you should be able to do the following.

\begin{itemize}
\item Identify limits that are of the indeterminate form $\frac{\#}{0}$.
\item Evaluate limits of the form $\frac{\#}{0}$.
\item Discuss what it means for a limit to equal $\infty$.
\item State the limit definition of a vertical asymptote.
\item Identify vertical asymptotes by looking at a graph.
\item Recognize when a limit is indicating there is a vertical asymptote.
\item Evaluate the limit as $x$ approaches a point where there is a vertical asymptote.
\item Find vertical asymptotes of common functions.
\item Match graphs of functions with their equations based on vertical asymptotes.
\item Find the limit as $x$ approaches $\pm \infty$ from a graph.
\item Calculate the limit as $x$ approaches $\pm \infty$ of common functions algebraically.
\item Define a horizontal asymptote.
\item Identify horizontal asymptotes by looking at a graph.
\item Find horizontal asymptotes using limits.
\item Recognize that a curve can cross a horizontal asymptote.
\item Produce a function with given asymptotic behavior.
\end{itemize}
\end{sectionOutcomes}

\phantom{text}%%% Making vertical spaces


%---Prereq notes/links----------%
%\textbf{Skills you may want to brush up on first} \\ To be ready to achieve these objectives, you may need to review the following trigonometry and algebra topics: 

%Eventually, maybe a rational functions xard will be created that would be helpful as a pre-req review?


\end{document}
