\documentclass{ximera}

\input{../preamble.tex}

\outcome{Define continuity in terms of limits.}
\outcome{Calculate limits using the limit laws.}
\outcome{Famous functions are continuous on their domains.}

\title[Dig-In:]{Continuity}
\begin{document}
\begin{abstract}
Continuity is defined by limits. 
\end{abstract}
\maketitle


Limits are simple to compute when they can be found by plugging the
value into the function.  That is, when
\[
\lim_{x\to c}f(x) = f(c).
\]
We call this property \textit{continuity}.

\begin{definition}
  A function $f$ is \dfn{continuous at a point} $a$ if
  \[
  \lim_{x\to a}f(x) = f(a).
  \]
\end{definition}

\begin{question}
  Consider the graph of $y=f(x)$ below
  \begin{image}
    \begin{tikzpicture}
      \begin{axis}
	[xmin=-0.2,
          xmax=2.2,
          ymin=-0.2,
          ymax=2.2,
          axis lines=center,
          xlabel=$x$,ylabel=$y$,
          every axis y label/.style={at=(current axis.above origin),anchor=south},
          every axis x label/.style={at=(current axis.right of origin),anchor=west},
	  domain=-1:2,
          clip=false,
	  ytick={0.5,1,1.5,2},
	  yticklabels={$0.5$,$1$,$1.5$,$2$},
	  xtick={0.5,1.0,1.5,2},
	  xticklabels={$0.5$,$1$,$1.5$,$2$},
	  grid = major
	]
        \addplot[very thick,penColor] plot coordinates {(0,0) (1,1)};
        \addplot[very thick,penColor] plot coordinates {(1,2) (2,0)};
        
        %\draw[very thin,color=black] (axis cs:0,-0.2) -- (axis cs:0,2);

        \addplot[color=penColor,fill=background,only marks,mark=*] coordinates{(1,1)};  %% open hole
        \addplot[color=penColor,fill=background,only marks,mark=*] coordinates{(.5,.5)};  %% open hole

        \addplot[color=penColor,fill=penColor,only marks,mark=*] coordinates{(0,0)};  %% closed hole
        \addplot[color=penColor,fill=penColor,only marks,mark=*] coordinates{(2,0)};  %% closed hole
        \addplot[color=penColor,fill=penColor,only marks,mark=*] coordinates{(1,2)};  %% closed hole
        
        %% \draw[fill=black] (axis cs:0,0) circle [radius=2pt];
	%% \draw[fill=black] (axis cs:2,0) circle [radius=2pt];
        %% \draw[fill=black] (axis cs:1,2) circle [radius=2pt];
	
	\end{axis}
    \end{tikzpicture}
  \end{image}
  Which of the following are true?
  \begin{multipleChoice}
    \choice{$f$ is continuous at $x=0.5$}
    \choice{$f$ is continuous at $x=1$}
    \choice[correct]{$f$ is continuous at $x=1.5$}
  \end{multipleChoice}
  
\end{question}

It is very important to note that saying
\begin{center}
  ``a function $f$ is continuous at a point $a$''
\end{center}
is really making \textbf{three} statements:
\begin{enumerate}
\item $f(a)$ is defined.  That is, $a$ is in the domain of $f$.
\item $\displaystyle\lim_{x\to a} f(x)$ exists.
\item $\displaystyle\lim_{x\to a} f(x) = f(a)$.
\end{enumerate}

The first two of these statements are implied by the third statement.

\begin{example}
Find the discontinuities (the values of $x$ where the function is not continuous) for the function described below:
\begin{image}
\begin{tikzpicture}
	\begin{axis}[
            domain=0:10,
            ymax=5,
            ymin=0,
            %samples=100,
            axis lines =middle, xlabel=$x$, ylabel=$y$,
            every axis y label/.style={at=(current axis.above origin),anchor=south},
            every axis x label/.style={at=(current axis.right of origin),anchor=west},
            %% ytick={0.5,1,1.5,2},
	    %% yticklabels={$0.5$,$1$,$1.5$,$2$},
	    %% xtick={0.5,1.0,1.5,2},
	    %% xticklabels={$0.5$,$1$,$1.5$,$2$},
	    grid = major
          ]
	  \addplot [very thick, penColor, smooth, domain=(4:10)] {3 + sin(deg(x*2))/(x-1)};
          \addplot [very thick, penColor, smooth, domain=(0:4)] {1};
          \addplot[color=penColor,fill=background,only marks,mark=*] coordinates{(4,3.30)};  %% open hole
          \addplot[color=penColor,fill=background,only marks,mark=*] coordinates{(2,1)};  %% open hole
          \addplot[color=penColor,fill=background,only marks,mark=*] coordinates{(6,2.893)};  %% open hole
          \addplot[color=penColor,fill=penColor,only marks,mark=*] coordinates{(4,1)};  %% closed hole
          \addplot[color=penColor,fill=penColor,only marks,mark=*] coordinates{(6,2)};  %% closed hole

          \addplot[color=penColor,fill=penColor,only marks,mark=*] coordinates{(0,1)};  %% closed hole
          \addplot[color=penColor,fill=penColor,only marks,mark=*] coordinates{(10,3.1)};  %% closed hole
        \end{axis}
\end{tikzpicture}
%% \caption{A plot of a function with discontinuities at $x=4$ and $x=6$.}
%% \label{plot:discontinuous-function}
\end{image}

\begin{explanation}
  To start, $f$ is not even defined at $x=\answer[given]{2}$, hence $f$
  cannot be continuous at $x=\answer[given]{2}$.

  Next, from the plot above we see that $\displaystyle\lim_{x\to 4} f(x)$ does not
  exist because
  \[
  \lim_{x\to 4^-}f(x) = \answer[given]{1}\qquad\text{and}\qquad \lim_{x\to 4^+}f(x) \approx \answer[given,tolerance=1]{3.5}
  \]
Hence $\lim_{x\to 4} f(x) \ne f(4)$, and so $f(x)$ is not
continuous at $x=4$.

We also see that $\displaystyle\lim_{x\to 6} f(x) \approx \answer[given,tolerance=.5]{3}$ while $f(6) =
\answer[given]{2}$. Hence $\displaystyle\lim_{x\to 6} f(x)$ does not exist, and so $f$ is not
continuous at $x=6$.
\end{explanation}
\end{example}


Building from the definition of \textit{continuity at a point}, we can
now define what it means for a function to be \textit{continuous} on
an open interval.

\begin{definition}
  A function $f$ is \dfn{continuous on an open interval} $I$ if
  $\displaystyle\lim_{x\to a} f(x) = f(a)$ for all $a$ in $I$.
\end{definition}

Loosely speaking, a function is continuous on an interval $I$ if you
can draw the function on that interval without any breaks in the
graph.  This is often referred to as being able to draw the graph
``without picking up your pencil.''

\begin{theorem}[Continuity of Famous Functions]\index{continuity of famous functions}\label{theorem:continuity}
The following functions are continuous on the given intervals for $k$ a real number and $b$ a positive real number:
\begin{description}
\item[Constant function]\index{constant function} $f(x) =k$ is continuous on $-\infty < x < \infty$.
\item[Identity function]\index{indenity function} $f(x) = x$ is continuous on $-\infty < x < \infty$.
\item[Power function]\index{power function} $f(x)=x^b$ is continuous on $-\infty < x < \infty$.
%\item[Exponential function]\index{exponential function} $f(x)=b^x$ is continuous on $-\infty < x < \infty$.
%\item[Logarithmic function]\index{logarithmic function} $f(x)=\log_b(x)$ is continuous on $0 < x < \infty$.
\item[Sine and cosine] Both $\sin(x)$ and $\cos(x)$ are continuous on $-\infty < x < \infty$.\index{sine}\index{cosine}
\end{description}
In essence, we are saying that the functions listed above are
continuous wherever they are defined, that is, on their natural
domains.
\end{theorem}


\begin{question}
  Compute:
  $\displaystyle\lim_{x\to 3} x^\pi \begin{prompt}= \answer{3^\pi}\end{prompt}$
  \begin{feedback}
    The function $f(x)=x^\pi$ is of the form $x^k$ for a real number
    $k$.  Therefore, $f(x)=x^\pi$ is continuous for all real values of
    $x$.  In particular, $f(x)$ is continuous at $x=3$.  Since $x^\pi$
    is continuous at $3$, we know that $\displaystyle\lim_{x\to 3} f(x) = f(3)$.
    That is, $\displaystyle\lim_{x\to 3} x^\pi = 3^\pi$
  \end{feedback}  
\end{question}

\begin{question}
  True or false: If $f$ and $g$ are continuous functions on an
  interval $I$, then $f\pm g$ is continuous on $I$.
  \begin{multipleChoice}
    \choice[correct]{True}
    \choice{False}
  \end{multipleChoice}
  \begin{feedback}
    This follows from the Sum/Difference Law.
  \end{feedback}
\end{question}

\begin{question}
  True or false: If $f$ and $g$ are continuous functions on an
  interval $I$, then $f/g$ is continuous on $I$.
  \begin{multipleChoice}
    \choice{True}
    \choice[correct]{False}
  \end{multipleChoice}
  \begin{feedback}
    In this case, $f/g$ will not be continuous for $x$ where $g(x) =
    0$.
  \end{feedback}
\end{question}


\begin{theorem}[Continuity of Polynomial Functions]\index{continuity of polynomial functions}
  All polynomial functions, meaning functions of the form
  \[
  f(x) = a_nx^n + a_{n-1}x^{n-1} + \dots + a_1 x + a_0
  \]
  where $n$ is a whole number and each $a_i$ is a real number, are
  continuous for all real numbers.
\end{theorem}

\begin{theorem}[Continuity of Rational Functions]\index{continuity of rational functions}
  Let $f$ and $g$ be polynomials.  Then a rational function, meaning an
  expression of the form
  \[
  h=\frac{f}{g}
  \]
  is continuous for all real numbers except where $g(x)=0$.  That is,
  rational functions are continuous wherever they are defined.
\begin{explanation}
      Let $a$ be a real number such that $g(a)\neq 0$.  Then, since
      $g(x)$ is continuous at $a$, $\displaystyle\lim_{x\to a} g(x) \neq 0$.
      Therefore, write with me, 
      \[
      \displaystyle\lim_{x \to a} h(x) = \displaystyle\lim_{x\to a} \frac{f(x)}{g(x)}
      \]
      and now by the Quotient Law, 
      \[
      \frac{\displaystyle\lim_{x\to a} f(x)}{\displaystyle \lim_{x\to a} g(x)}
      \]
      and by the continuity of polynomials we may now set $x=a$
      \[
      \frac{f(a)}{g(a)}=h(a).
      \]
      Since we have shown that $\displaystyle\lim_{x\to a} h(x) = h(a)$, we have
      shown that $h$ is continuous at $x=a$.
\end{explanation}
\end{theorem}

\begin{question}
  Where is $f(x) = \frac{x^2-3x+2}{x-2}$ continuous?
  \begin{multipleChoice}
    \choice{for all real numbers}
    \choice{at $x=2$}
    \choice[correct]{for all real numbers, except $x=2$}
    \choice{impossible to say}
  \end{multipleChoice}
\end{question}


Now, we give basic rules for how limits interact with composition
of functions.

\begin{theorem}[Composition Limit Law]\index{composition limit law}
  If $f(x)$ is continuous at $x = \displaystyle\lim_{x\to a} g(x)$, then
  \[
  \displaystyle\lim_{x\to a} f(g(x)) = f(\displaystyle\lim_{x\to a} g(x)).
  \]
\end{theorem}

Because the limit of a continuous function is the same as the function
value, we can now pass limits inside continuous functions.

\begin{corollary}[Continuity of Composite Functions]
If $g$ is continuous at $x=a$, then $f(g(x))$ is continuous at $x=a$.
\end{corollary}

\begin{example}
  Compute the following limit using limit laws:
  \[
  \displaystyle\lim_{x \to 0} \sqrt{\cos(x)}
  \]
  \begin{explanation}
    By continuity of $x^k$, assuming $\displaystyle\lim_{x \to 0} \cos(x) >0$,
    \[
    \displaystyle\lim_{x \to 0} \sqrt{\cos(x)} = \sqrt{\displaystyle\lim_{x \to 0} \cos(x)},
    \]
    and now since cosine is continuous for all real numbers,
    \[
    \sqrt{\cos(0)} = \sqrt{1} = 1.
    \]
  \end{explanation}
\end{example}

%%LEFT AND RIGHT CONTINUITY
\section{Left and right continuity}


At this point we have a small problem.  For functions such as
$\sqrt{x}$, the natural domain is $0\leq x <\infty$.  This is not an
open interval.  What does it mean to say that $\sqrt{x}$ is continuous
at $0$ when $\sqrt{x}$ is not defined for $x<0$? To get us out of this
quagmire, we need a new definition:

\begin{definition}
  A function $f$ is \dfn{left continuous} at a point $a$ if
  $\displaystyle\lim_{x\to a^-} f(x) = f(a)$.

  A function $f$ is \dfn{right continuous} at a point $a$ if
  $\displaystyle\lim_{x\to a^+} f(x) = f(a)$.
\end{definition}

Now we can say that a function is continuous at a left endpoint of an
interval if it is right continuous there, and a function is continuous
at the right endpoint of an interval if it is left continuous
there. This allows us to talk about continuity on closed intervals.

\begin{question}
Here we give the graph of a function defined on $[0,10]$.
\begin{image}
\begin{tikzpicture}
	\begin{axis}[
            domain=0:10,
            ymax=5,
            ymin=0,
            samples=100,
            axis lines =middle, xlabel=$x$, ylabel=$y$,
            every axis y label/.style={at=(current axis.above origin),anchor=south},
            every axis x label/.style={at=(current axis.right of origin),anchor=west}
          ]
	  \addplot [very thick, penColor, smooth, domain=(4:10)] {3 + sin(deg(x*2))/(x-1)};
          \addplot [very thick, penColor, smooth, domain=(0:4)] {1};
          \addplot[color=penColor,fill=background,only marks,mark=*] coordinates{(4,3.30)};  %% open hole
          \addplot[color=penColor,fill=background,only marks,mark=*] coordinates{(6,2.893)};  %% open hole
          \addplot[color=penColor,fill=penColor,only marks,mark=*] coordinates{(4,1)};  %% closed hole
          \addplot[color=penColor,fill=penColor,only marks,mark=*] coordinates{(6,2)};  %% closed hole
          \addplot[color=penColor,fill=penColor,only marks,mark=*] coordinates{(0,1)};  %% closed hole
          \addplot[color=penColor,fill=penColor,only marks,mark=*] coordinates{(10,3.1)};  %% closed hole
        \end{axis}
\end{tikzpicture}
%% \caption{A plot of a function with discontinuities at $x=4$ and $x=6$.}
%% \label{plot:discontinuous-function}
\end{image}
What are the largest intervals of continuity for this function?
\begin{multipleChoice}
  \choice{$[0,10]$}
  \choice{$[0,4)$ and $(4,10]$}
  \choice{$[0,4]$, $[4,6]$, and $[6,10]$}
  \choice{$(0,4)$, $(4,6)$, and $(6,10)$}
  \choice{$[0,4]$, $(4,6)$, and $[6,10]$}
  \choice[correct]{$[0,4]$, $(4,6)$, and $(6,10]$}
  \choice{$[0,4)$, $(4,6)$, and $(6,10]$}
  \choice{$(0,4]$, $[4,6]$, and $[6,10)$}
\end{multipleChoice}
\begin{feedback}
Notice that our function is left continuous at $x=4$ so we can include
$4$ in the interval $[0,4]$. Four is not included in the interval
$(4,6)$ because our function is not right continuous at $x=4$.
Similarly, our function is neither right or left continuous at $x=6$,
so $6$ is not included in any intervals.  Our function is left
continuous at $x=0$ and right continuous at $x=10$ so we included
these endpoints in our intervals.
\end{feedback}
\end{question}



\end{document}

