\documentclass{ximera}

\input{../../../../preamble.tex}

\title{1.1 Practice Quiz}

\begin{document}
\begin{abstract}
Answer the questions honestly to assess your own abilities. This survey will not be graded. It can be used for you to determine your own strengths and weaknesses.
\end{abstract}
\maketitle

\begin{problem} 
    Could you factor the following polynomial?
    $$2x^2 - 4$$
  \begin{multipleChoice}
      \choice[correct]{Yes, I can factor this polynomial.}
      \choice[correct]{Yes, but I would need to review some resources first.}
      \choice[correct]{No, I would need to learn/relearn how to factor.}
      \choice[correct]{I don't know.}
      \choice[correct]{I don't want to answer this question right now.}
  \end{multipleChoice}
\end{problem}
        

\begin{problem}
  Which of the following approximate the slope of the ``zoomed line''? Select all that are correct.
  \begin{selectAll}
    \choice{$\frac{(f(a)+h) - f(a)}{(a+h)-a}$}
    \choice[correct]{$\frac{f(a+h) - f(a)}{(a+h)-a}$}
    \choice{$\frac{(f(a)-h) - f(a)}{(a-h)-a}$}
    \choice[correct]{$\frac{f(a-h) - f(a)}{(a-h)-a}$}
    \choice{$\frac{f(a) - (f(a)+h)}{a-(a+h)}$}
    \choice[correct]{$\frac{f(a) - f(a+h)}{a-(a+h)}$}
    \choice{$\frac{f(a) - (f(a)-h)}{a-(a-h)}$}
    \choice[correct]{$\frac{f(a) - f(a-h)}{a-(a-h)}$}
  \end{selectAll}
\end{problem}

\begin{problem}
   Let $f(x) = 3x-1$.  Zoom in on the curve around $a = -2$ so that $h
   = 0.1$.  Use one of the formulations in the problem above to
   approximate the slope of the curve.  The slope of the curve at $a =
   -2$ is approximately\dots \begin{prompt}$\answer{3}$\end{prompt}
\end{problem}

\begin{problem}
   Repeat the previous problem for $f(x) = x^2 - 1$, $a = 0$, and $h =
   0.2$.  Choose a formulation that will give you a positive answer
   for the slope.  The (positive) slope of the curve at $a = 0$ is
   approximately\dots\begin{prompt} $\answer{0.2}$\end{prompt}
\end{problem}


\begin{problem}
   Zoom in on the curve $f(x) = x^2 - 1$ near $x=0$ again.  By looking
   at the graph, what is your best guess for the actual slope of the
   curve at zero?
   \begin{multipleChoice}
     \choice{impossible to say}
     \choice[correct]{zero}
     \choice{one}
     \choice{infinity}
   \end{multipleChoice}
\end{problem}

%\input{../leveledQuestions.tex}


\end{document}
