\documentclass[handout]{ximera}
%\documentclass[10pt,handout,twocolumn,twoside,wordchoicegiven]{xercises}
%\documentclass[10pt,handout,twocolumn,twoside,wordchoicegiven]{xourse}


%\license{Creative Commons 3.0 By-NC}
\input{../preamble.tex}

\outcome{Practice Optimization.}
   

\title[Exercises:]{Optimization Exercises}

\begin{document}
\begin{abstract}
  Here we'll practice solving optimization problems.
\end{abstract}
\maketitle

%%OPTIMIZATION 1
\begin{exercise}
  Find the dimensions of the rectangle of largest area having fixed
  perimeter $100$.

  \[
  \answer{25}\times \answer{25}
  \]

\end{exercise}

%%OPTIMIZATION 2
\begin{exercise}
 Find the dimensions of the rectangle of smallest perimeter with fixed area $100$.
  \[
  \answer{10}\times \answer{10}
  \]

\end{exercise}

%%OPTIMIZATION 3
\begin{exercise}
A box with square base and no top is to hold a volume $100$.  Find
  the dimensions of the box that requires the least material for the
  five sides.

  \[
  \text{width}=\answer{2\cdot 5^{2/3}},\qquad
  \text{length}=\answer{2\cdot 5^{2/3}},\qquad
  \text{height}=\answer{5^{2/3}}
  \]

\end{exercise}

%%OPTIMIZATION 4
\begin{exercise}
A box with square base is to hold a volume $200$. The bottom and top
  are formed by folding in flaps from all four sides, so that the
  bottom and top consist of two layers of cardboard.  Find the
  dimensions of the box that requires the least material.

  \[
  \text{width}=\answer{100^{1/3}},\qquad
  \text{length}=\answer{100^{1/3}},\qquad
  \text{height}=\answer{2\cdot 100^{1/3}}
  \]

\end{exercise}

%%OPTIMIZATION 5
\begin{exercise}
A box with square base and no top is to hold a volume $V$.  Find (in
  terms of $V$) the dimensions of the box that requires the least
  material for the five sides.

  \[
  \text{width}=\answer{2^{1/3}V^{1/3}},\qquad
  \text{length}=\answer{2^{1/3}V^{1/3}},\qquad
  \text{height}=\answer{V^{1/3}/2^{2/3}}
  \]

\end{exercise}

%%OPTIMIZATION 6
\begin{exercise}
You have $100$ feet of fence to make a rectangular play area
  alongside the wall of your house.  The wall of the house bounds one
  side.  What is the largest size possible (in square feet) for the
  play area?

  \[
  \answer{1250} square feet
  \]

\end{exercise}


%%OPTIMIZATION 7
\begin{exercise}
Marketing tells you that if you set the price of an item at \$10
  then you will be unable to sell it, but that you can sell 500 items
  for each dollar below \$10 that you set the price.  Suppose your
  fixed costs total \$3000, and your marginal cost is \$2 per item.
  What is the most profit you can make?
  \begin{prompt}
  \[
  \answer{5000}
  \]
  \end{prompt}
\end{exercise}

%%OPTIMIZATION 8
\begin{exercise}
Find the area of the largest rectangle that fits inside a semicircle
  of radius $10$ (one side of the rectangle is along the diameter of
  the semicircle).

  \[
  \answer{100}
  \]

\end{exercise}

%%OPTIMIZATION 16
\begin{exercise}
A rectangle is inscribed in the ellipse
  \[
  \frac{x^2}{3}+\frac{y^2}{5}=1
  \]
  Find the dimensions of the rectangle with the largest area.

  \begin{align*}
  \text{base} &= \answer{\sqrt{6}}\\
  \text{height} &= \answer{\sqrt{10}}
  \end{align*}

\end{exercise}

%%OPTIMIZATION 12
\begin{exercise}
What fraction of the volume of a sphere is taken up by the largest cylinder
  that can be fit inside the sphere?

  \[
  \answer{1/\sqrt{3}}
  \]

\end{exercise}


%%OPTIMIZATION 15
\begin{exercise}
A trough is constructed with the following dimensions:
  \begin{image}
    \begin{tikzpicture}[scale=.5]
      \draw[penColor, very thick] (-2,4) -- (2,3.5) -- (0,0)--cycle;
      \draw[penColor, very thick] (0,0) -- (9,1) -- (11,4.5) --(2,3.5);
      \draw[penColor,very thick] (-2,4) -- (7,5) -- (7.45,4.11);
      \draw[penColor,very thick] (11,4.5) -- (7,5);
      \draw[penColor,very thick,dashed] (7.45,4.11) -- (9,1);
      \node at (9,5.1) {$x$};
      \node at (3,5) {$64$};
      \node at (-1.5,2) {$2$};
      \node at (1.5,1.8) {$2$};
    \end{tikzpicture}
  \end{image}
  Find $x$ that maximizes the volume.

  \[
  x = \answer{2\sqrt(2)}
  \]

\end{exercise}


%%OPTIMIZATION 19
\begin{exercise}
Find the area of the rectangle with the largest area that fits
  beneath the curve $y=\sqrt{x}$ on the interval $[0,4]$.

    \[
    A = \answer{16/(3\cdot \sqrt{3})}
    \]

\end{exercise}














\end{document}
